---
title: "Grocery associate rule mining"
author: "Soumik Choudhuri"
output:
  word_document: default
  pdf_document: default
---
# ASSOCIATE RULE MINING


## Treatment of the data

In this step, we first took the Groceries dataset where each row represented the items bought by a user
We then created a dataframe to have only 2 columns - user_id and product where user_id is the row number for groceries datasets
and grouped the products by user_id

```{r ans f-1, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}

library(tidyverse)
library(arules)  # has a big ecosystem of packages built around it
library(arulesViz)
library(reshape2)

temp=read.csv('groceries.txt',header=FALSE)
temp=rownames_to_column(temp)
tt=melt(temp,id.vars = 'rowname')
tt= tt[tt$value!='',]

tt=tt[,c('rowname','value')]
colnames(tt)=c('user_id','product')

tt$user_id=strtoi(tt$user_id)
tt=tt[order(tt$user_id),]

tt$user_id=factor(tt$user_id)
prod_lt_by_user = split(x=tt$product, f=tt$user_id)
prod_lt_by_user = lapply(prod_lt_by_user, unique)

```


## Summary of the products distribution

```{r ans f-2, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}

prod_dist = as(prod_lt_by_user, "transactions")
summary(prod_dist)

```


## OBSERVATIONS

In the summary of the prod_dist, we find that "whole milk" was the highest bought item followed by "other vegetables"
Also, we found that in majority (around 7079 times), 4 items were bought in a single transaction.


## Running the apriori function to get the support, confidence and lift

```{r ans f-3, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}

prod_for_users = apriori(prod_dist, 
                     parameter=list(support=.005, confidence=.1, maxlen=5))

plot(prod_for_users)
```


## OBSERVATIONS

From the scatter plot we found that there are a few items with support less than 0.05 but have a very high confidence above 0.3
which indicated that there is a good association between these items
So we decided to further inspect these items.



## Inspection steps and Observations

We started with the inspection of product combination with confidence above 0.3

```{r ans f-4, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}

inspect(subset(prod_for_users, confidence > 0.3))
```


Upon inspection, we found that "onions" and "other vegetables" have the highest lift indicating a strong tendency of other vegetables being bought with onions
Same is the case with "root vegetables" and "other vegetables" or vegetables along with milk which have lift close to 3
Among these, we find the highest count of 625 for the combination of "other vegetables" and "whole milk".


Next, we inspect the product combinations with count > 150 and lift > 1.5

```{r ans f-5, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}

inspect(subset(prod_for_users, count > 150 & lift > 1.5 ))
```

Upon putting the above condition of count and lift, we find that "pip fruit" and "tropical fruit" have a lift close to 4 indicating high association between them


We find that people generally buy whole milk and other vegetables together upon inspecting for lift > 1 and support > 0.03

```{r ans f-6, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}
inspect(subset(prod_for_users, subset=lift > 1 & support > 0.03))
```


Upon putting the condition of support > 0.10, we find that these are values for individual items like "soda" or "rolls/buns"

```{r ans f-7, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}
inspect(subset(prod_for_users, subset = support > 0.10))
```



## NETWORK GRAPHS


We then plot the network graph to see the association between the different items bought.

```{r ans f-8, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}
sub1 = subset(prod_for_users, subset=confidence > 0.01 & support > 0.005)
plot(head(sub1, 100, by='lift'), method='graph')

```

## OBSERVATIONS

1)We get to see that other vegetables and whole milk are the items which are highest associated with other items i.e. they have the highest network.

2) Also, we find bottled water, beer, soda, chocolate and juice belong to a completely different cluster indicating they are not associated with vegetables or whole milk or other regular items



## GEPHI PLOTS

Finally, we use the Gephi software to visualize the association between the items.

```{r ans f-9, include=TRUE, echo=FALSE, fig.height=10, fig.width=15}

#install.packages("BiocManager") 
#BiocManager::install("EBImage")
library(EBImage)
img = readImage("Groceries.png")
display(img, method = "raster")
img2 = readImage("grocery_color_degree_size_betweeness.png")
display(img2, method = "raster")
```


## OBSERVATIONS

1) Here we see that "other vegetables" and "whole milk" have the biggest font followed by "root vegetables", "yogurt", "sausage" and others.

2) The font of an item is proportional to the degree of the node which in turn is related to the association of that item.
The Gephi plot also validates the earlier network plot indicating that vegetables and whole milk have highest association and are often bought with other items

3) In the second Gephi plot we find that there are 2 clusters validating our earlier inference that bottled water, beer, soda are not associated with the regular items and are mostly bought separately.

